---
title: "aggreCAT: an R Package for Mathematically Aggregating Expert Judgements"
format:
    jss-pdf:
      keep-tex: true
      header-includes:
        - \usepackage[utf8]{inputenc}
        - \usepackage{amsmath}
        - \usepackage{amsfonts}
        - \usepackage{caption}
        - \usepackage{booktabs}
        - \usepackage{longtable}
        - \usepackage{array}
        - \usepackage{multirow}
        - \usepackage{wrapfig}
        - \usepackage{float}
        #- \usepackage{colortbl}
        - \usepackage{pdflscape}
        - \usepackage{tabu}
        - \usepackage{threeparttable}
        - \usepackage{threeparttablex}
        - \usepackage[normalem]{ulem}
        - \usepackage{makecell}
        - \usepackage{xcolor}
        - \newcommand{\blandscape}{\begin{landscape}}
        - \newcommand{\elandscape}{\end{landscape}}
        - \usepackage{underscore}
        # - \usepackage[authoryear,round]{natbib}
      x11names: true
      journal:
        cite-shortnames: true
        type: article
    jss-html:
        number-depth: 3
author:
    # use this syntax to add text on several lines
    # To add another line, use \AND at the end of the previous one as above
  - name: 'Elliot Gould^[School of Ecosystem and Forest Sciences, University of Melbourne]'
    email: 'elliot.gould (at) unimelb.edu.au'
    orcid: '0000-0002-6585-538X'
    affiliations: 'University of Melbourne'
    address: |
      | School of Forest and Ecosystem Sciences
      | University of Melbourne, Parkville, Victoria 3010
  - name: "Charles T. Gray"
    affiliations: 'Newcastle University'
  - name: "Aaron Willcox"
    affiliations:  'University of Melbourne'
    orcid: '0000-0003-2536-2596'
  - name: "Rose O'Dea"
    affiliations: 'University of New South Wales'
    orcid: '0000-0001-8177-5075'
  - name: "Rebecca Groenewegen"
    affiliations:  'University of Melbourne'
    orcid: '0000-0001-9177-8536'
  - name: 'David P. Wilkinson'
    orcid: '0000-0002-9560-6499'
    affiliations: 'University of Melbourne'
abstract: |
  Structured elicitation protocols, such as the IDEA protocol, may be used to elicit expert judgements in the form of subjective probabilities from multiple experts. Judgements from individual experts about a particular phenomena must therefore be mathematically aggregated into a single prediction. The process of aggregation may be complicated when judgements are elicited with uncertainty bounds, and also when there are several rounds of elicitation. This paper presents the new R package [aggreCAT]{.pkg}, which provides 29 unique aggregation methods for combining individual judgements into a single, probabilistic measure. The aggregation methods were developed as a part of the Defense Advanced Research Projects Agency (DARPA) 'Systematizing Confidence in Open Research and Evidence' (SCORE) programme, which aims to generate confidence scores or estimates of 'claim credibility' for over 4000 research claims from the social and behavioural sciences. We provide several worked examples illustrating the underlying mechanics of the aggregation methods. We also describe a general workflow for using the software in practice to facilitate uptake of this software for appropriate use-cases.
keywords: [mathematical aggregation, expert judgement, DARPA SCORE, replicability, R]
keywords-formatted: [mathematical aggregation, expert judgement, DARPA SCORE, replicability, "[R]{.proglang}"]
bibliography: bibliography.bib
editor:
  markdown:
    wrap: 72
callout-appearance: simple
callout-icon: false
appendix-style: none
# pdf-engine: citeproc #until quarto repsects classoption:shortnames
---

```{r, setup, include=FALSE, message=FALSE}
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE, comment = "#>")
options(tinytex.verbose = TRUE)
options(width = 80)
suppressPackageStartupMessages(library(tidyverse))
library(aggreCAT)
library(tinytex)
library(knitr)
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
library(tinytable)
```

# Introduction {#sec-introduction}

Expert judgement is frequently used to inform forecasting about
uncertain future events across a range of fields and applications
including ecological management and decision making [@Legge2022],
biosecurity risk analyses [@Wittmann2015], replicability forecasts
[@Mody2026], and horizon scanning exercises [@Sutherland2017].
Judgements from groups of experts tend to perform better than a single
expert [@Goossens2008], and it is best-practice to elicit judgements
from diverse groups so that group members can bring "different
perspectives, cross-examine each others' reasoning, and share
information", however judgements or forecasts must then be distilled
into a single forecast, ideally accompanied by estimates of uncertainty
around those estimates [@Hanea2021]. Judgements from multiple experts
may be combined into a single score using either behavioural approaches
that force experts into forming consensus, or by using mathematical
approaches [@Goossens2008].

Aggregated elicited judgements form a critical component of
decision-making across multiple contexts, and while there are a variety
of methods for mathematically aggregating expert judgements into single
point-predictions, there has been a dearth of accessible tools for
implementing any aggregation more complex than linear averages. Many
existing software tools are closed-source, are irreproducible, or
written in programming languages rarely used by ecologists or
metascientists. Methods implemented in R, one of the most widely used
programming languages in the life sciences [@Gao2025], are limited in
scope, or have been archived from CRAN. For instance, the [R]{.proglang}
[@R] package [expert]{.pkg} provides three models of expert opinion to
combine judgements elicited from groups of experts (CITE) , and
[SHELF]{.pkg} implements only a single method (weighted linear pool) for
aggregating expert judgements (CITE). Other [R]{.proglang} packages
providing methods to mathematically aggregate expert judgements do so
for non-point predictions, for example, [opera]{.pkg}, which generates
time-series predictions (CITE).

aggreCAT is an open, reliable, and modular R package for the
mathematical aggregation of expert elicitation data with a
straightforward user interface. The aggreCAT package was developed by
the by the repliCATS (Collaborative Assessment for Trustworthy Science)
project [@Fraser:2021] as part of the DARPA SCORE (Systematizing
Confidence in Open Research and Evidence) program [@alipourfard2021],
which aimed to generate quantitative 'Confidence Scores' — estimates of
the likely replicability of research claims from the social and
behavioural sciences — for use as a proxy measure of credibility in the
absence of direct replication effort. aggreCAT arose from the need for a
software solution to reliably conduct mathematical aggregation at scale.
The repliCATS project used aggreCAT to aggregate expert elicited
judgements into Confidence Scores for $> 4000$ claims [@Mody2026]. While
originally a bespoke implementation for repliCATS, we have since
developed aggreCAT into a more general R package that can handle most
types of elicitation data requiring mathematical aggregation, allowing
for wider applicability.

The repliCATS project developed 29 different methods for mathematically
aggregating judgements elicited from groups of experts into Confidence
Scores [@Hanea2021]. We developed the [aggreCAT]{.pkg} package to
implement these aggregation methods and deliver Confidence Scores for
over 4000 SBS research claims as a part of the DARPA SCORE project.

aggreCAT provides a suite of 29 aggregation methods that explore
different approaches to mathematical aggregation from straightforward
arithmetic calculations to Bayesian statistical models. In addition, we
provide extra functionality such as plotting functions and performance
evaluation against known outcomes. While the repliCATS project uses the
IDEA protocol to structure their elicitation process (CROSS-REF), we
have generalised aggreCAT to work with a variety of elicitation
protocols. aggreCAT fills a large void in open software for aggregating
elicited judgements, providing enormous benefit to researchers and
decision makers across any field of research.

In this paper we give an overview of the [aggreCAT]{.pkg} package so
that researchers may apply any of the the aggregation functions
described in [@Hanea2021] and provided by aggreCAT to their own expert
elicitation datasets where mathematical aggregation is required.

# Mathematically Aggregating Expert Judgements

The aggregation methods in [aggreCAT]{.pkg} were developed in the
context of the IDEA protocol (CROSSREF Box 1), a structured elicitation
procedure that draws on the 'wisdom of crowds' to elicit probabilistic
judgements from groups of experts about uncertain events.

This elicitation structure directly determines which aggregation methods
in [aggreCAT]{.pkg} are applicable to a given dataset: most methods
require only the post-discussion (*Round 2*) three-point judgements,
some additionally use the pre-discussion (*Round 1*) judgements, and a
small number require supplementary ratings or externally collected data.
These requirements are summarised for each method in
@tbl-method-summary-table.

Although the aggregation methods were developed in the context of the
IDEA protocol, all methods requiring only a single round of elicitation
can be applied to any structured elicitation procedure that generates
lower, upper, and best estimates from each individual without enforcing
behavioural consensus.

:::: callout
::: {#ideaProtocol}
### Box 1: The repliCATS IDEA Protocol {.unnumbered}

The repliCATS project adapted and deployed the IDEA protocol to elicit
crowd-sourced judgements from diverse groups about the likely
replicability of Social and Behavioural Science (SBS) research claims
[@Fraser:2021]. The IDEA ('Investigate', 'Discuss', 'Estimate' and
'Aggregate') protocol is a four-step structured elicitation protocol
that draws on the 'wisdom of crowds' to elicit subjective judgements
about the likelihood of uncertain events [@hemming2017, figure 1]. To
collect expert judgements about the replicability of SBS claims, we
asked participants to estimate the "probability that direct replications
of a study would find a statistically significant effect in the same
direction as the original claim", eliciting estimates of uncertainty in
the form of upper and lower bounds on those point-estimates.

The IDEA protocol structures elicitation across two rounds: individuals
first provide judgements independently (*Round 1*), before anonymously
reviewing one another's estimates and engaging in group discussion,
after which they provide a revised set of judgements (*Round 2*). In
each round, participants provide probabilistic three-point judgements —
a best estimate and upper and lower bounds — alongside supplementary
ratings of plausibility, comprehensibility, and involvement in the
original study. Thus, for a single assessment, 2 sets of judgements are
elicited from each expert, *pre-* and *post-*group discussion, referred
to here on as *Round 1* and *Round 2* judgements, respectively.

During the *Investigate* phase, individuals review the claim and draw on
background knowledge to provide numeric estimates of the percent
probability of the claim replicating.

These estimates are numeric (0 - 100%) and include a best estimate, and
upper and lower bounds on that estimate.

<!-- Not sure whether to include the below or not? Can it be explained elsewhere? -->

Individuals also provide a value on the likert binary scale up to 7,
rating the individuals' degree of comprehension of the claim.
Individuals were also asked to provide short comments justifying their
estimates.

In the *Discuss* phase, three-point estimates from each group member are
anonymously presented to the group, who then collectively discuss
differences in opinion and provide potential evidence for these
differences.

Next, in the *Estimate* phase, individuals are asked to provide a second
set of three-point probablistic judgements.

During the fourth phase, *Aggregate*, judgements are mathematically
aggregated into a single Confidence Score, or in the repliCATS case,
forecast of replicability.

![The IDEA protocol as deployed by the repliCATS project (reproduced
with permission from Wintle et al.
2021).](images/img_IDEA_repliCATS.png "The IDEA protocol as deployed by the repliCATS project (reproduced with permission from Wintle et al. 2021"){#fig1}
:::
::::

## Notation and Problem Formulation

We describe some preliminary mathematical notation used to represent
each aggregation method. The total number of research claims, $claim$,
or unique forecasts being assessed, $C$ , is indexed by $c = 1, ..., C$.
For each claim $c$, an individual $i$ assesses the probability of a
claim replicating by providing three probabilities: a lower bound
${L}_{i,c}$, an upper bound ${U}_{i,c}$, and a best estimate $B_{i,c}$,
satisfying the inequalities: $0 \le Li,c \le Bi,c \le Ui,c \le 1$. Every
claim $c$ is assessed by multiple experts, with the total total number
of individuals is denoted by $N$, and is indexed by $i = 1, ..., N$.
These probabilities are aggregated using one of the aggregation methods
to obtain a group or aggregate probability, denoted by $\hat{p}_c$. The
aggregated probability calculated using a specific method is given by
$\hat{p}_{c}\left(Method \space ID \right)$. Each aggregation is
assigned a unique $Method \space ID$ which is the abbreviation of the
mathematical operation used in calculating the weights. Note that all
Best, Lower and Upper estimates are taken to be `round 2` judgements
from the repliCATS IDEA protocol [Figure 1](#fig1), Box 1), unless
appended by a "1", where they are `round 1` judgements, e.g. $B1_{i,c}$
denotes the `round 1` Best estimate from individual $i$ for claim $c$.

Equal-weighting of judgements are less calibrated, accurate and
informative than weighted aggregation methods where judgements from
experts who performed well in similar judgement tasks are more heavily
weighted [@Hanea2021]. Proxies for forecasting performance, such as
breadth and variability of qualitative reasons used by experts to
justify their judgements, can be used to form weights in the absence of
measures of experts' prior performance [@Hanea2021].

The aggregation methods other than the [AverageWAgg]{.fct} and Bayesian
approaches in [aggreCAT]{.pkg} each employ weighting schemes that are
informed by proxies for good forecasting performance whereby experts'
estimates are weighted differently by measures of reasoning, engagement,
openness to changing their mind in light of new facts, evidence or
opinions presented in the discussion round, extremity of estimates,
informativeness of estimates, asymmetry of estimate bounds, granularity
of estimates, and by prior statistical knowledge as measured in a quiz.

Below, we define standardised notation for describing weighted linear
combinations of individual judgements where un-normalised weights are
denoted by $w\_method$ and normalised weights by $\tilde{w} \_ method$
(@eq-eqn1). Given that for all aggregation methods weights are
normalised, and that the normalisation process is the same for each
aggregation method, the equations for the aggregation methods are
presented for un-normalised weights.

$$\hat{p}_c\left(Method \space ID \right) = \frac{1}{N}\sum_{i=1}^N \tilde{w}\_{method}_{i,c}  B_{i,c}$$

Each claim *outcome* (i.e. the outcome of a replication study) assumes
binary values, where the value is 0 if the claim is false, and 1 if the
claim is true. '`TRUE`' claims are claims where the replication study
found a significant result in the same direction as the original
research claim, and '`FALSE`' claims are those where the replication
study *did not* find a significant result in the same direction as the
original study.

By default, weights are calculated across all claims on a
per-individual, per-claim basis, such that judgements for the same
individual are weighted differently across all claims they have provided
judgements for. There are some exceptions to this default: `GranWAgg`,
`QuizWAgg`, `IndIntWAgg` `IndIntAsymWAgg`, `VarIndIntWAgg`,
`KitchSinkWAgg`. Note that `IndIntWAgg`, and methods that include its
weighting function `weight_nIndivInterval()` as a component, re-scale
weights by a fixed measure across all claims. Hence, for aggregation
methods that use information from multiple claims other than the target
claim for which the Confidence Score is being computed, each individual
claim $c$ is indexed by $d = 1, ..., C$. Where the default weighting is
used, this is coded into each function. However, where more complex and
method-specific weighting methods are used, modularised functions have
been created for ease of debugging. These function names are prefixed
with `weight_`.

## Aggregation Wrapper Functions

Although there are 29 aggregation methods in total, we grouped methods
based on their mathematical properties into eight 'wrapper' functions,
denoted by the suffix `WAgg`, the abbreviation of *weighted
aggregation*: `LinearWAgg()`, `AverageWAgg()`, `BayesianWAgg()`,
`IntervalWAgg()`, `ShiftingWAgg()`, `ReasoningWAgg()`,
`DistributionWAgg()`, and `ExtremisationWAgg()`. The specific
aggregation *method* is applied according to the `type` argument, whose
options are described in each aggregation wrapper functions' help page.

### 'Tidy' Aggregation and Required Inputs

The design philosophy of [aggreCAT]{.pkg} is principled on 'tidy' data
[@Wickham:2014vp]. Each aggregation method expects a
[data.frame]{.class} or [tibble]{.class} of judgements as its input, and
returns a [tibble]{.class} containing the variables `method`,
`paper_id`, `cs` and `n_experts` (see @sec-AverageWAgg for illustration
of outputs); where `method` is a character vector corresponding to the
aggregation method name specified in the `type` argument. Each
aggregation is applied as a summary function [@Wickham2017R], and
therefore returns a single row with a single confidence score `cs` for
each claim or `paper_id`. The number of expert judgements summarised in
the aggregated confidence score is returned in `n_experts`. Because of
the tidy nature of the aggregation outputs, multiple aggregation methods
can be applied to the same data with the results of all aggregation
methods row-bound together in a single `tibble` (see the example
repliCATS workflow in @sec-workflow).

Every aggregation function requires at minimum the three-point
probabilistic judgements (`question == "direct_replication"`), however
some methods require only best estimates
(`element == "three_point_best"`), and some require multiple rounds of
elicitation. Only the aggregation method *CompWAgg* additionally
requires comprehensibility ratings (`question == "comprehension"`). For
a full summary of each aggregation method, its calling function, and
data requirements, see @tbl-method-summary-table.

## Package Datasets

The [aggreCAT]{.pkg} package includes several datasets collected by the
repliCATS project during a pilot study at a two-day workshop in the
Netherlands in July 2019, at which 25 participants assessed the
replicability of 25 unique SBS research claims with known outcomes using
the IDEA protocol [@Hanea2021].

The core dataset `data_ratings` is a tidy [data.frame]{.class} of
probabilistic three-point judgements (lower bound, best estimate, upper
bound) elicited from participants for each claim, across two rounds of
elicitation. Additional datasets [@Fraser:2021] capture participants'
written justifications (`data_justifications`), peer comments and
ratings (`data_comments`), and three supplementary datasets collected
externally to the IDEA protocol: a statistical knowledge quiz
(`data_supp_quiz`), qualitatively coded reasoning categories
(`data_supp_reasons`), and model-derived Bayesian priors
(`data_supp_priors`). Known replication outcomes sourced from previous
large-scale replication projects [@Klein2014; @Klein2018ManyL2;
@Ebersole2016; @Camerer2018; @aac4716] for a subset of claims are
provided in `data_outcomes`.

Full descriptions of each dataset, including variable definitions, data
structure, and example data, are provided in the [aggreCAT]{.pkg}
package documentation (`?data_ratings`) and in the Package Datasets
vignette available on the [aggreCAT]{.pkg} pkgdown website.

# Demonstrating Mathematical Aggregation with aggreCAT

We now demonstrate how judgements elicited from a diverse group of
individuals may be mathematically aggregated for a single forecasting
problem, using the datasets provided by [aggreCAT]{.pkg}. We illustrate
the internal mechanics of the weighting methods and the different data
requirements of each of the different types of aggregators -- namely;
methods with non-weighted linear combinations of judgements, weighted
linear combinations of judgements, re-scaled weighted linear
combinations of judgements, methods that require supplementary data, and
methods that require data elicited from the full IDEA protocol. Each
group of methods differs in the type of judgements elicited (single- or
three-point estimates), the number of elicitation rounds (one or two
rounds), whether multiple forecasts / elicited judgements are used
during confidence score computation for a target forecast / claim, and
finally whether supplementary data is required for aggregation.

## Focal Claim Aggregation {#sec-focal-claims}

Here we demonstrate the application of aggregation methods for each
group of methods using a set of 'focal claims' selected from the pilot
study dataset supplied with the [aggreCAT]{.pkg} package. Below we
subset the dataset `data_ratings` to include a sample of four claims
with judgements from five randomly-sampled participants. From these
focal claims, we select a target claim for which we will apply an
exemplar aggregation method from each mathematical aggregator
(@tbl-focal-claim ).

```{r}
#| label: focal-claim-selection
#| prompt: true
set.seed(1234)
focal_claims <- data_ratings %>%
  dplyr::filter(paper_id %in% c("24", "138", "186", "108"))
# select 5 users to highlight in focal claim demonstration
focal_users <- focal_claims %>%
  dplyr::distinct(user_name) %>%
  dplyr::slice_sample(n = 5)
# filter out non-focal users from focal claims
focal_claims <- focal_claims %>%
  dplyr::right_join(focal_users, by = "user_name")
focal_claims
```

`focal_claims` contains 220 rows across four claims and five
participants. Each row records a single `element` of a participant's
assessment for a given `paper_id` and `round`. The `question` column
distinguishes the three-point probabilistic judgements about
replicability (`direct_replication`, with `element` values
`three_point_lower`, `three_point_best`, and `three_point_upper`) from
supplementary assessment data: participants' plausibility ratings
(`belief_binary`), comprehensibility ratings (`comprehension`), and
involvement in the original study (`involved_binary`). All aggregation
functions require at minimum the `direct_replication` rows, and expect
data in this long, tidy format.

We will work with data from a subset of participants from a single claim
`108` for a subset of participants who assessed this claim
(@tbl-focal-claim).

```{r}
#| label: tbl-focal-claim
#| tbl-cap: "Focal Claim Data: Round 2 expert judgements for claim 108 derived from a subset of 5 claims and 5 participants from `data\_ratings`. Judgements are displayed as percentages."
#| eval: true
#| echo: false
#| results: asis
focal_claims %>%
  filter(question == "direct_replication",
         paper_id == "108",
         round == "round_2") %>%
  pivot_wider(names_from = element,
              values_from = value) %>%
  arrange(user_name) %>%
  select(paper_id,
         user_name,
         three_point_lower,
         three_point_best,
         three_point_upper) %>%
  rename(`Claim ID` = paper_id,
         `User Name` = user_name,
         `Lower Bound` = three_point_lower,
         `Best Estimate` = three_point_best,
         `Upper Bound` = three_point_upper) %>%
  tt()
```

# References
