---
title: "aggreCAT datasets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{aggreCAT datasets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: vignette_references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
library(aggreCAT)
library(tidyverse)
```


The [aggreCAT]{.pkg} package includes the core dataset `data_ratings`
consisting of judgements elicited during a pilot experiment exploring
the performance of IDEA groups in assessing replicability of a set of
claims with "known outcomes." "Known-outcome" claims are SBS research
claims that have been subject to replication studies in previous
large-scale replication projects[^1]. Data were collected using the
repliCATS IDEA protocol at a two day workshop[^2] in the Netherlands, on
July 2019, at which 25 participants assessed the replicability of 25
unique SBS claims. In addition to the probabilistic estimates provided
for each research claim assessed, participants were also asked to rate
the claim's plausibility and comprehensibility, answer whether they were
involved in any aspect of the original study, and to provide their
reasoning in support of their quantitative estimates, which were used to
form measures of reasoning breadth and engagement [@Fraser:2021].

[^1]: Many labs 1, 2 and 3 @Klein2014, @Klein2018ManyL2, @Ebersole2016,
    the Social Sciences Replication Project @Camerer2018 and the
    Reproducibility Project Psychology @aac4716.

[^2]: See @Hanea2021 for details. The workshop was held at the annual
    meeting of the Society for the Improvement of Psychological Science
    (SIPS), [\<https://osf.io/ndzpt/\>](https://osf.io/ndzpt/){.uri}.

## Formatted Judgement Data

`data_ratings` is a *tidy* [data.frame]{.class} wherein each
*observation* (or row) corresponds to a single value in the set of
`value`s constituting a participant's complete assessment of a research
claim. Each research claim is assigned a unique `paper_id`, and each
participant has a unique (and anonymous) `user_name`. The variable
`round` denotes the round in which each `value` was elicited (`round_1`
or `round_2`). `question` denotes the type of question the `value`
pertains to; `direct_replication` for probabilistic judgements about the
replicability of the claim, `belief_binary` for participants' belief in
the plausibility of the claim, `comprehension` for participants'
comprehensibility ratings, and `involved_binary` for involvement in the
original study. An additional column `element` maintains the tidy
structure of the data, while capturing the multiple `value`s that
comprise a full assessment of the replicability (`direct_replication`)
of a claim; `three_point_best`, `three_point_lower` and
`three_point_upper` denote the best estimate and lower and upper bounds
respectively. `binary_question` describes the `element` for both the
plausibility rating (`belief_binary`) and involvement
(`involved_binary`) questions, whereas `likert_binary` is the `element`
describing a participant's `comprehension` rating. Judgements are
recorded in column `value` in the form of percentage probabilities
ranging from (0,100). The `binary_question`s corresponding to
comprehensibility and involvement consist of binary values (`1` for the
affirmative, and `-1` for the negative). Finally, values corresponding
to participants' comprehension ratings are on a `likert_binary` scale
from `1` through `7`. Note that additional columns with participant
attributes can be included in the ratings dataset if required by the
user; we include the `group` column in `data-ratings`, which describes
the group number the participant was a part of. Below we show some
example data for a single user for a single claim to illustrate this
structure of the core `data_ratings` dataset.

```{r data_ratings-sample, message=TRUE, results='hold'}
aggreCAT::data_ratings %>% 
  dplyr::filter(paper_id == dplyr::first(paper_id), 
                user_name == dplyr::first(user_name)) %>% 
  head()

```

Not all data necessary for constructing weights on performance is
contained in `data_ratings`. Additional data collected as part of the
repliCATS IDEA protocol are contained within separate datasets to
`data_ratings`. Participants provided justifications for giving
particular judgemetns, and these are contained in `data_justifications`.
On the repliCATS platform users were given the option to comment on
others' justifications (`data_comments`), to vote on others' comments
(`data_comment_ratings`) and on others' justifications
(`data_justification_ratings`). Finally, [aggreCAT]{.pkg} contains three
'supplementary' datasets containing data collected externally to the
repliCATS IDEA protocol: `data_supp_quiz`, `data_supp_priors`, and
`data_supp_reasons`.

## Quiz Score Data {#sec-quiz-supplementary-data}

Prior to the workshop, participants were asked to complete an optional
quiz on statistical concepts and meta-research which we thought would
aid in reliably evaluating the replicability of research claims. Quiz
responses are contained in `data_supp_quiz` and are used to construct
performance weights for the aggregation method `QuizWAgg`, where each
participant receives a `quiz_score` if they completed the quiz, and `NA`
if they did not attempt the quiz [see @Hanea2021 for further details]. Additional methods of scoring the quiz responses are provided in `data_supp_quiz`. 

```{r data_supp_quiz-sample, message=TRUE, results='hold'}
aggreCAT::data_supp_quiz
```


## Reasoning Data {#sec-reasonwagg-supplementary-data}

The `ReasonWAgg` aggregation type uses the number of unique reasons given by participants to
support a Best Estimate for a given claim $B_{i,c}$ to construct
performance weights, and is contained within `data_supp_reasons`.
Qualitative statements made by individuals during claim evaluation were
recorded on the repliCATS platform [@Pearson2021] and coded as falling
into one of 25 unique reasoning categories by the repliCATS Reasoning
team [@Wintle:2021]. Reasoning categories include plausibility of the
claim, effect size, sample size, presence of a power analysis,
transparency of reporting, and journal reporting [@Hanea2021]. Within
`data_supp_reasons`, each of the reasoning categories that passed our
inter-coder reliability threshold are distributed as columns in the
dataset whose names are prefixed with `RW`, and for each claim
`paper_id`, each participant `user_id` is assigned a logical `1` or `0`
if they included that reasoning category in support of their Best
estimate for that claim. See `ReasoningWAgg()` for details on the
`ReasonWAgg` aggregation method.

```{r data_supp_reasons-sample}
aggreCAT::data_supp_reasons %>% 
  glimpse()
```


## Bayesian Prior Data {#sec-bayesian-supplementary-data}

The method `BayPRIORsAgg` (implemented in `BayesianWAgg()`) uses Bayesian updating to update a prior
probability of a claim replicating estimated from a predictive model
[@Gould2021a] using an aggregate of the best estimates for all
participants assessing a given claim $c$ [@Hanea2021]. The prior data is
contained in `data_supp_priors` with each claim in column `paper_id`
being assigned a prior probability (on the logit scale) of the claim
replicating in column `prior_means`.

```{r data_supp_priors-sample}
aggreCAT::data_supp_priors
```


# TODO

- [ ] `data_comments`
- [ ] `data_confidence_scores`
- [ ] `data_justifications`
- [ ] `data_outcomes`

# References